Master's Project Journal

### 2018-11-13 tex_files/Journal.tex

I then took a sample measurement of the clock output of the ring oscillator running at 5 MHz and began to process it in Matlab to get an idea of the jitter. To do this I wrote a basic script which calculated the mean and standard deviation of the periods in the sample as well and the Time Interval Error.
This week I also met with Eugene and demonstrated my code for him as well as giving him binary files containing 8 MSamp of data for both the ring oscillator and the phase accumulator. He will be able to analyse their jitter characteristics and from there begin to determine the optimum control coefficients.\\
%TODO pierre or filip?
Worked also with Philipe getting his oscillator to work, using a similar design to mine however implemented in VHDL.\\

@@

\section*{Phase Detection}
The next step then was to work on a phase detection mechanism for the ADPLL. First off I went down the simplest route and implemented a 1 bit bang-bang detector using just a single D flip flop. After meeting with Brian on Wednesday we discussed some of the challenges of implementing the phase detector on an FPGA. Foremost among these was the issue of metastability and the likelihood of both signals occurring at the same time. For an implementation on an ASIC where an arbitration circuit can be constructed using gates I do not have this luxury, as I am restricted to simple look-up table elements. As a temporary workaround for this we decided I would implement a phase detector with the assumption that the signals are synchronised to the FPGA clock.\\
Working under this assumption I created a phase detector which counts the number of FPGA clock cycles between edges of the two signals in question. In order to avoid issues where the rising edge of the second signal detected occurs very shortly before the next rising edge of the first signal detected the maximum phase difference that can be detected is the high time of the first signal.\\
This phase detector outputs signed 2s complement integers that can be fed directly into the loop filter which succeeds this block.

@@

>>>>

I then took a sample measurement of the clock output of the ring oscillator running at 5 MHz and began to process it in Matlab to get an idea of the jitter. To do this I wrote a basic script which calculated the mean and standard deviation of the periods in the sample as well and the Time Interval Error.
This week I also met with Eugene and demonstrated my code for him as well as giving him binary files containing 8 MSamp of data for both the ring oscillator and the phase accumulator. He will be able to analyse their jitter characteristics and from there begin to determine the optimum control coefficients.\\
Worked also with Pierre getting his oscillator to work, using a similar design to mine however implemented in VHDL.\\

@@

\section*{Phase Detection I}
The next step then was to work on a phase detection mechanism for the ADPLL. First off I went down the simplest route and implemented a 1 bit bang-bang detector using just a single D flip flop. After meeting with Brian on Wednesday we discussed some of the challenges of implementing the phase detector on an FPGA. Foremost among these was the issue of metastability and the likelihood of both signals occurring at the same time. For an implementation on an ASIC where an arbitration circuit can be constructed using gates I do not have this luxury, as I am restricted to simple look-up table elements. As a temporary workaround for this we decided I would implement a phase detector with the assumption that the signals are synchronised to the FPGA clock.\\
Working under this assumption I created a phase detector which counts the number of FPGA clock cycles between edges of the two signals in question. In order to avoid issues where the rising edge of the second signal detected occurs very shortly before the next rising edge of the first signal detected the maximum phase difference that can be detected is the high time of the first signal.\\
This phase detector outputs signed 2s complement integers that can be fed directly into the loop filter which succeeds this block.

@@

\section*{Phase Detection II}
The above phase detector was created using a state machine with the following diagram describing its behaviour:\\
%TODO insert image of the state transitions
It is preceded by two synchronisers which attempt to avoid metastable conditions, which will be used for the time being so that I could advance until we found a method that worked more generally.\\
I was able to test this phase detector and confirm that it was measuring a frequency difference in line with what I expected given my simulation results, despite spending too much time trying to figure out why the display was not giving me a reasonable output when it was just being updated far too quickly to be human readable.\\
%TODO insert image of the RTL diagram
I then realised I was also operating at a 160 MHz clock in both simulation and hardware so I attempted to push the clock up to the 400 MHz mark then but started to run into so timing issues on both inter and intra clock paths so I had to dial back the FPGA clock frequency to 258 MHz which was enough to remove all but what I think is an allowable timing violation on a synchroniser.\\
Working with the 1:0.1 rule for proportional/integral gains I then implemented a PI loop filter as in Eldar's PhD on Friday in order to complete an entire multi-bit ADPLL. After ironing out one issue caused by a lack of bit extension causing strange behaviour on the control code if it was negative which had gone undetected in simulation I was able to get my PLL locking to an on-chip reference signal late on Friday. In more detail the 8 bit signed output of the error loop filter was being added to a 12 bit bias signal and due to a lack of sign extension -1 became a +511. This error would correct itself on the next cycle leading to signal which looked like: $||--|||||-----||--|||||$\\%TODO do I have a screenshot of this error?
Pierre told me about his method for the arbiter so maybe I can implement that method myself rather than relying on my limited use case version.
%TODO insert a good simulation
